\pagestyle{myheadings} \setcounter{page}{1} \setcounter{footnote}{0}

\section{~To do list (tickets)} \label{app:to_do}
\newcounters 
\vssub

\noindent
Item numbers will eventually become the Trac ticket numbers \ldots

\begin{itemize}

\item[95] Many: keep list of upgrades up to data in intro/about.tex.

\item[96] Many: make group list in intro/group.tex complete and up to
  date. Will need to put out an E-mail about who should be on here.

\item[97] Tolman: Final version of Acknowledgements with fixed date.

% \item[98] DONE. Tolman: Provide documentation for GMD in eqs/NL3.tex.

% \item[99] DONE. Tolman: Provide documentation for nonlinear filter in
%  eqs/NLS.tex.

% \item[---] Banner / Morrison: Will ST5 go into distribution? Answer is no.

\item[70] Zieger: Provide documentation for ST6.

\item[---] ???: What happened with source term BT2. Will this go into
  distribution version?

\item[---] ???: What happened with source term BT3. Will this go into
  distribution version?

\item[---] Tolman to find out from Ardhuin: Should reflection be in the source
  terms or in the propagation part. Right now they are in both. Please
  review. If it goes into source term part, we might want to make a new switch
  group. It is now in both eqs/unknown.tex and num/shore.tex. 

\item[100]: Alves: Check one more time list of outputs. There is some text at
  the end of eqs/output.tex that I commented out. Please review and deal with
  as needed.

\item[---] Tolman / Rogers / Ardhuin: Does the present description of the
  spatial schemes suffice. See Erick's remarks in SCRIP appendix.

\item[101] Li: Provide documentation for spatial UNO scheme in
  num/space\_trad\_2.tex. Follow format and level of detail from UP scheme in
  num/space\_trad\_3.

\item[102] Li: Provide documentation for spatial SMC grid in num/space\_SMC.tex.

\item[103] Li: Provide documentation for rotated grid option in
  num/rotagrid.tex.

\item[104] Li: Provide documentation for spectral UNO scheme in
  num/spec\_uno.tex.

\item[105] Roland / Ardhuin: Add dynamic time stepping documentation for
  refraction in num/spec.tex?

\item[---] Van der Westhuysen: Check up num/part.tex and num/track.tex,
  particularly the ID tables at the head of the sections. Are the references to
  XWaves etc. still up to date, or should we mention IMEDS now?

% \item[106] ???: Provide documentation for ww3\_bounc in run/ww3bounc.tex.

% \item[107] Ardhuin: Provide documentation for ww3\_prtide in run/ww3prtide.tex.

% \item[108] DONE: Tolman: Provide documentation for ww3\_gspl in run/ww3gspl.tex.

\item[109] Chawla: Provide documentation for ww3\_gint in run/ww3gint.tex.

\item[100] Alves: Make sure field output is consistent in all inp files.

\item[110] Chawla: Check and update tables at the end of Chapter 4 (run.tex)

\item[---] ???: Final sanity check on run/ww3\_grid.tex

\item[---] ???: Final sanity check on run/ww3\_strt.tex

\item[---] ???: Final sanity check on run/ww3\_bound.tex

\item[---] ???: Final sanity check on run/ww3\_bounc.tex

\item[---] ???: Final sanity check on run/ww3\_prep.tex

\item[---] ???: Final sanity check on run/ww3\_prnc.tex

\item[---] ???: Final sanity check on run/ww3\_prtide.tex

\item[---] ???: Final sanity check on run/ww3\_shel.tex

\item[---] ???: Final sanity check on run/ww3\_gspl.tex

\item[---] ???: Final sanity check on run/ww3\_multi.tex

\item[---] ???: Final sanity check on run/ww3\_gint.tex

\item[---] ???: Final sanity check on run/ww3\_outf.tex

\item[---] ???: Final sanity check on run/ww3\_ounf.tex

\item[---] ???: Final sanity check on run/gx\_outf.tex

\item[---] ???: Final sanity check on run/ww3\_grib.tex

\item[---] ???: Final sanity check on run/ww3\_outp.tex

\item[---] ???: Final sanity check on run/ww3\_ounp.tex

\item[---] ???: Final sanity check on run/gx\_outp.tex

\item[---] ???: Final sanity check on run/ww3\_trck.tex

\item[---] ???: Final sanity check on run/ww3\_systrk.tex

\item[111] Alves: Update tar files for tar installation, list of directories
  created and document svn install scripts in bin directory list. File
  impl/install.tex. 

\item[112] Rogers: document SCRIP switches in impl/switch.tex.

\item[113] Ardhuin: document TIDE and REFT switches in impl/switch.tex.

\item[114] Chawla: Orphaned XYG and LLG switches in code. Find or document.

\item[115] Alves / Rogers: add regtests and cases to end of impl/test\_it.tex.

\item[70] Zieger: Add w3snl6md.ftn to sys/files\_w3.tex

\item[116] Ardhuin: Add w3sbs1md.ftn missing documentation in sys/files\_w3.tex

\item[---] Ardhuin: What to do with w3reflmd.ftn in sys/files\_w3.tex.

\item[117] Tolman: Update figures in sys/files\_w3.tex.

\item[118] Tolman: Add figures to sys/files\_wm.tex.

\item[119] Tolman: Update figures in sys/aux.tex.

% \item[120] Tolman: Check all of sys/vars.tex.

\item[---] All: Check documentation in data structure files.

\item[---] Tolman: Final editing for release version of manual.

\end{itemize}

\bpage
\pagestyle{empty}
