\vsssub
\subsubsection{~$S_{ice}$: Damping by sea ice (Shen et al.)} \label{sec:ICE3}
\vsssub

\opthead{IC3}{Clarkson U. Fortran-77 code}{E. Rogers, X. Zhao, S. Cheng, S. Zieger}

\noindent
The third method for representing wave-ice interaction is taken from\linebreak
\cite{art:WS10}. This model treats the ice as a visco-elastic
layer. ${C_{ice,1}}$ is used for ice thickness (m); ${C_{ice,2}}$ is used for
the viscosity ($\mathrm{m^2\,s^{-1}}$); ${C_{ice,3}}$ is used for density
($\mathrm{kg\,m^{-3}}$); ${C_{ice,4}}$ is used for effective shear modulus
(Pa); ${C_{ice,5}}$ is not used. An example setting is 
${C_{ice,1...4}}=[0.1, 1.0, 917.0, 0.0]$.

In \ws\ version 4, this method of $S_{ice}$ ({\code IC3}) was much more expensive than {\code IC1} or {\code IC2}. This issue is largely addressed in model version 5.

The namelist {\F SIC3} is introduced in model version 5. The namelist parameters are summarized in a list here, and some are discussed in further detail below.

\begin{clist}
\cit {IC3CHENG} {Solution technique new in version 5. Default = {\code TRUE}.}
\cit {IC3HILIM} {Optional limiter on ice thickness. Default=100 (i.e. by default, the option is not used).}
\cit {IC3KILIM} {Optional limiter on dissipation rate ${k_i}$. Default=100 (i.e. by default, the option is not used).}
\cit {USECGICE} {When set to {\code TRUE}, the model will include the effect of ice on the group velocity. Default = {\code FALSE}.}
\cit {IC3VISC} {If user wishes to use an effective viscosity that is constant and uniform, this can now be done via namelist. Default=N/A.}
\cit {IC3ELAS} {As with {\code IC3VISC}, but for effective elasticity. Default=N/A.}
\cit {IC3DENS} {As with {\code IC3VISC}, but for ice density. Default=N/A.}
\cit {IC3HICE} {As with {\code IC3VISC}, but for ice thickness. Default=N/A.}
\cit {IC3MAXCNC} {Parameter which can be used to optionally switch to another dissipation for some ice conditions (see below). Default=100 (i.e. option is not used). Normal range is 0 to 1.}
\cit {IC3MAXTHK} {Idem. Default=100 (i.e. option is not used). Normal range is 0 to 10 meters.}        
\cit {IC2REYNOLDS} {Parameter associated with IC2 non-dispersive turbulent boundary layer scheme. Default=1.5e+5.}
\cit {IC2ROUGH} {Idem. Default=0.02.}
\cit {IC2SMOOTH} {Idem. Default=7.0e+4.}
\cit {IC2VISC} {Idem. Default=2.0.}
\cit {IC2TURB} {Idem. Default=2.0.}          
\cit {IC2TURBS} {Idem. Default=0.0.}
\end{clist}

The {\code IC3CHENG} option is new in model version 5. When set to {\code TRUE}, the model will use an alternative solution technique provided by S. Cheng. This has two important features. First, stability is improved, such that there is no need to use the ice limiter, i.e. the {\code IC3HILIM} parameter. Second, this method requires that three of four ice rheology parameters be stationary and uniform, input via namelist parameters (see below).

If {\code IC3CHENG} is set to {\code FALSE}, the user is advised to use the ice thickness limiter {\code IC3HILIM} to ensure stability (value of 25 to 100 cm is suggested). The parameter {\code IC3KILIM} was required for stable and fast computations in some prior development versions of \ws, but is now unnecessary and may be ignored by the user.

In model version 4.18, four ice rheology parameters (ice thickness, effective viscosity, effective elasticity, and ice density) were allowed to be nonstationary and non-uniform. This could be provided using {\file ww3\_prep}. Or in cases where {\file ww3\_shel} is used and non-uniform input is unnecessary, the ``homogeneous'' option of {\file ww3\_shel} was available for rheology input. In model version 5, an option is added to specify the four ice rheology parameters via the namelist {\F SIC3}. Two restrictions apply: 1) If {\code IC3CHENG} is set to {\code FALSE} and  {\code USECGICE} is set to {\code TRUE}, the namelist method cannot be used, and 2) If {\code IC3CHENG} is set to {\code TRUE}, the namelist method must be used for three of the rheology parameters (effective viscosity, effective elasticity, and ice density). If {\code IC3CHENG} is set to {\code TRUE} or {\code USECGICE} is set to {\code FALSE}, the fourth ice rheology parameter (ice thickness) can be input by either method (namelist or non-namelist). The model performs error checking to ensure that the user has specified input for each parameter by a single method (neither method of input is assumed to supercede the other).

The ${k_r}$ modified by ice is incorporated into the governing
equation~(\ref{eq:bal_plane}) via the $c_g$ (group velocity) and $c$ (phase velocity) calculations on the
left-hand side; e.g. \citet[][and subsequent unpublished work]{art:RH09}.  The
modified wavenumber ${k_r}$ produces effects analogous to shoaling and
refraction by bathymetry. To activate the shoaling effect, the model should be
operated with namelist variable {\code USECGICE = TRUE}. To activate the refraction effect, the model should be
compiled with switch {\code REFRX}. With this switch, the model computes
refraction based on spatial gradients in phase velocity that include ice effects, rather than the
simpler wave dispersion relation without ice. These effects are demonstrated in the regression test {\file
ww3\_tic1.3} which is provided with the code.

The group velocity using {\code IC3CHENG} solver with zero ice thickness does not collapse exactly to that from the open water dispersion relation. [This is caused by numerical error in the calculation $c_g = \frac{\partial \sigma}{\partial k} = \frac{\Delta \sigma}{\Delta k} $. This calculation will be improved in a subsequent patch.] These small differences in group speed will result in slight shoaling and refraction errors if these effects are turned on. Error for ice thickness=zero was found to be less than 10\% and was frequency dependent. This has been avoided by skipping the solver if ice thickness is exactly zero. If ice thickness is close to but not exactly zero, then the issue may persist. The solutions from CHENG for other parameters (effective viscosity, effective shear modulus) as they approach zero were not tested. Small but material difference have also noted between the solutions from {\code IC3CHENG} set to {\code FALSE} vs. {\code TRUE} for the same ice inputs.

As noted above, {\code USECGICE = TRUE} is required for the shoaling effect. However, since some ice rheology will lead to an increase in group velocity, the user is advised to be careful with this option. The group velocity affects the CFL criterion, which may require that the user reduce the time step size. {\code USECGICE = FALSE} is recommended for users that do not wish to worry about this issue.

In model version 5, a non-default option is added which causes the dissipation parameterization to change for some ice conditions. If ice concentration exceeds {\code IC3MAXCNC} and ice thickness exceeds {\code IC3MAXTHK}, the {\code IC2} dissipation (more specifically, the non-default, non-dispersive boundary layer scheme sub-option of {\code IC2}) is used in place of the dispersion-based dissipation estimate of \cite{art:WS10}. See description of {\code IC2} for more information. Since it is non-dispersive, this feature should not be used with {\code USECGICE = TRUE}.
