\vsssub
\subsubsection{~$S_{ice}$: Damping by sea ice (effective medium models)} \label{sec:ICE5}
\vsssub

\opthead{IC5}{U. of Otago MATLAB code}{Q. Liu, E. Rogers, A. Babanin}

\noindent
The fifth option ({\code IC5}) for representing ice-induced wave decay was introduced by \citet{Liu2020}. It provides three different models to estimate the attenuation rate $k_i$, including two viscoelastic (VE) beam models \citep{art:MMS15} and one viscous model \citep{Meylan2018}.

Each of the two VE models, i.e., the extended Fox and Squire model \citep[hereafter the EFS model; see also][]{art:FS1994} and the Robinson and Palmer model \citep[][hereafter the RP model]{Robinson1990}, describes the infinitely-long, floating ice layer as a homogeneous VE medium, the characteristics of which are represented by two \emph{empirical} rheological parameters, namely the elastic shear modulus of the ice layer $G$ and the viscosity of the ice layer $\eta$. Under this paradigm, the dispersion relation is modified as \citep{art:MMS15}:
\begin{equation}
Q g \kappa \tanh(\kappa d)-\sigma^2 = 0,
\label{eq:vedisp}
\end{equation}
where $d$ is the water depth and $\kappa = k_r + i k_i$ is the complex wavenumber of ice-coupled waves. The real part $k_r = 2\pi / \lambda$ characterizes the effect of ice on the wavelength and $k_i$ is the attenuation rate of wave amplitude. The $Q$ term in (\ref{eq:vedisp}) for the EFS and RP models reads

\begin{align}
Q_{EFS} &= \frac{G_{\eta} h_i^3}{6 \rho_w g} (1+\nu) \kappa^4 - \frac{\rho_i h_i \sigma^2}{\rho_w g} + 1,\label{eq:fsdisp}\\
%
Q_{RP}  &= \frac{G h_i^3}{6 \rho_w g} (1+\nu) \kappa^4 - \frac{\rho_i h_i \sigma^2}{\rho_w g} + 1 - i \frac{\sigma \eta}{\rho_w g}.\label{eq:rpdisp}
\end{align}
Here $\rho_w$ ($\rho_i$) is the density of water (sea ice), $h_i$ is the ice cover thickness, $G_{\eta}=G- i \sigma \rho_i \eta$ is the complex shear modulus and $\nu \simeq 0.3$ is the Poisson ratio of sea ice. The viscosity parameter $\eta$ for the EFS model has the dimension of the kinematic viscosity (unit: m$^2$ s$^{-1}$), whereas $\eta$ for the RP model denotes a constant viscous damping force per unit area and per unit velocity \citep[unit: kg m$^{-2}$ s$^{-1}$;][]{Meylan2018}. Due to their high similarity, (\ref{eq:fsdisp}) and (\ref{eq:rpdisp}) can be computed by a single solver using the Newton-Raphson iterative method \citep[see Appendix of][]{Liu2020}.

Through a detailed theoretical analysis, \citet{Meylan2018} suggested under the assumption that $k_r$ does not deviate from the open-water wavenumber $k_0$ significantly, and that the attenuation rate $k_i$ is weak, the two VE models described above predict
\begin{align}
k_i^{EFS} &\propto \eta h_i^3 \sigma^{11},\label{eq:fspw}\\ k_i^{RP} &\propto \frac{\eta}{\rho_w g^2} \sigma^3,\label{eq:rppw}
\end{align}
whereas previous field measurements \citep[e.g.,][]{Meylan2018, RMK21} support a power law $k_i \propto \sigma^n$, with $n$ between 2 and 4. Eqs.~(\ref{eq:fspw}) and (\ref{eq:rppw}) indicate at certain regimes (i.e., $k_r \approx k_0$ and low $k_i$), $k_i$ of the EFS model is too sensitive to wave frequency and $k_i$ of the RP model shows no dependence on ice thickness.

The third model included in the {\code IC5} module is based on the ``Model with Order 3 Power Law'' proposed by \citet[][their section 6.2; hereafter the M2 model]{Meylan2018}, which assumes the loss of wave energy is proportional to the horizontal ice velocity squared times the ice thickness. The attenuation rate is given by
\begin{equation}
k_i^{M2} = \frac{\eta h_i}{\rho_w g^2} \sigma^3,
\label{eq:m2pw}
\end{equation}
where $k_i$ is linearly scaled with $h_i$, similar to the field measurements collected in \citet{art:Dob15} under pancake ice and $\eta$ is in kg m$^{-3}$ s$^{-1}$. 

Same as {\code IC3}, {\code IC5} requires four ice parameters as input: $C_{ice, 1}$ for ice thickness $h_i$ (m), $C_{ice, 2}$ for the \emph{effective} viscosity $\eta$ (unit varies with the model selected; see above), $C_{ice, 3}$ for ice density $\rho_i$ (kg m$^{-3}$) and $C_{ice, 4}$ for the \emph{effective} shear modulus $G$ (Pa)\footnote{This parameter is not relevant for the M2 model and one can simply use $G=1$ Pa when the M2 model is selected.}. Applications of the {\code IC5} module to two case studies were presented in \citet{Liu2020}. The reader is referred to this paper and the regression test {\code ww3\_tic1.1} for examples of how to use this ice module.

The namelist {\F SIC5} contains the parameters summarized below:
\begin{clist}
\cit{IC5MINIG} {the minimum allowed shear modulus $G$; Default= 1 Pa (i.e., zero $G$ is not allowed).}
%
\cit{IC5MINWT} {the minimum allowed wave periods $T$; Default=0 s (i.e., by default, this option is not used).}
%
\cit{IC5MAXKRATIO} {the maximum allowed $k_{ow}/k_r$, where $k_{ow}$ is the open-water wavenumber; Default=1E9 (i.e., by default, this option is not used).}
%
\cit{IC5MAXKI} {the maximum allowed $k_i$; Default=100 m$^{-1}$ (i.e., by default, this option is not used).}
%
\cit{IC5MINHW} {the minimum allowed water depth $d$; Default=0 m (i.e., by default, this option is not used).}
%
\cit{IC5MAXITER} {the maximum allowed \# of iteration; Default=100.}
%
\cit{IC5VEMOD} {the sea ice model to be selected: 1 - {\code EFS}, 2 - {\code RP}, 3 - {\code M2}; Default=3 (i.e., \textbf{the {\code M2} model is chosen}).}
\end{clist}
The first 6 parameters were introduced to improve the stability of the numerical solver for the EFS model \citep[the solver may fail for small wave periods in some rare cases, particularly for shallow water depth $d$ and low $G$; see][]{Liu2020}. Nonetheless, since version 7.12, the M2 model becomes the default option and these limiters are therefore not used by default.
