\vssub
\subsection{~Propagation}
\vssub

In a numerical model, a Eulerian form of the balance equation
(\ref{eq:balance0}) is needed. This balance equation can either be written in
the form of a transport equation (with velocities outside the derivatives), or
in a conservation form (with velocities inside the derivatives). The former
form is valid for the vector wavenumber spectrum $N(\bk;\bx,t)$ only, whereas
valid equations of the latter form can be derived for arbitrary spectral
formulations, as long as the corresponding Jacobian transformation as
described above is well behaved \citep[e.g.,][]{tol:GAOS98b}. Furthermore, the
conservation equation conserves total wave energy/action, unlike the transport
equation. This is an important feature of an equation when applied in a
numerical model. The balance equation for the spectrum $N(k,\theta;\bx,t)$ as
used in \ws\ is given as (for convenience of notation, the spectrum is
henceforth denoted simply as $N$):

%----------------------------------%
% Plain-grid propagation equations %
%----------------------------------%
% eq:bal_plane
% eq:x_dot
% eq:k_dot
% eq:theta_dot

\begin{equation}
\frac{\partial N}{\partial t} + \nabla_x \cdot {\bf\dot{x}}N +
\frac{\partial}{\partial k} \dot{k}N +
\frac{\partial}{\partial \theta} \dot{\theta}N =
\frac{S}{\sigma} \: , \label{eq:bal_plane}
\end{equation}
\begin{equation}
{\bf \dot{x}} = {\bf c}_g + {\bf U} \: , \label{eq:x_dot}
\end{equation}
\begin{equation}
\dot{k} = - \frac{\partial \sigma}{\partial d}
\frac{\partial d}{\partial s} - {\bk} \cdot
\frac{\partial {\bf U}}{\partial s} \: , \label{eq:k_dot}
\end{equation}
\begin{equation}
\dot{\theta} = - \frac{1}{k} \left [
\frac{\partial \sigma}{\partial d} \frac{\partial d}{\partial m}
+ {\bk} \cdot \frac{\partial {\bf U}}{\partial m}
\right ] \: , \label{eq:theta_dot}
\end{equation}

\noindent
where ${\bf c}_g = (c_g \sin \theta, c_g \cos \theta$, $s$ is a coordinate in the
direction $\theta$ and $m$ is a coordinate perpendicular to $s$.
Equation~(\ref{eq:bal_plane}) is valid for Cartesian coordinates. For large-scale
applications, this equation is usually transferred to spherical coordinates,
defined by longitude $\lambda$ and latitude $\phi$, but maintaining the
definition of the local variance \citep[i.e., per unit surface, as
in][]{art:WAM88}

%---------------------------------%
% Spherical propagation equations %
%---------------------------------%
% eq:bal_sphere
% eq:phi_dot
% eq:lambda_dot
% eq:theta_g_dot

\begin{equation}
\frac{\partial N}{\partial t} +
\frac{1}{\cos \phi} \frac{\partial}{\partial \phi} \dot{\phi}
    N \cos \theta +
\frac{\partial}{\partial \lambda} \dot{\lambda}N +
\frac{\partial}{\partial k} \dot{k} N +
\frac{\partial}{\partial \theta} \dot{\theta}_g N
= \frac{S}{\sigma} \: ,\label{eq:bal_sphere}
\end{equation}
\begin{equation}
\dot{\phi} = \frac{c_g \cos \theta + U_\phi}{R}
\: , \label{eq:phi_dot}
\end{equation}
\begin{equation}
\dot{\lambda} =  \frac{c_g \sin \theta + U_\lambda}{R \cos \phi}
\: , \label{eq:lambda_dot}
\end{equation}
\begin{equation}
\dot{\theta}_g = \dot{\theta} -
\frac{c_g \tan \phi \cos \theta}{R}
\: , \label{eq:theta_g_dot}
\end{equation}

\noindent
where $R$ is the radius of the earth and $U_\phi$ and $U_\lambda$ are current
components. Equation~(\ref{eq:theta_g_dot}) includes a correction term for
propagation along great circles, using a Cartesian definition of $\theta$
where $\theta = 0$ corresponds to waves traveling from west to east. \ws\ can
be run using either Cartesian or Spherical coordinates. Note that unresolved obstacles
such as islands can be included in the equations. In \ws\ this is done at the
level of the numerical scheme, as is discussed in section~\ref{sub:num_obst}. Also, depth variations 
at the scale of the wavelength can be introduced by a scattering source term described in 
 section~\ref{sec:BS1}. 
 
 Finally, both Cartesian  and spherical coordinates can be discretized in many ways, using quadrangles (rectangular, curvilinear or SMC grids) and triangles. 
That aspect is treated in chapter~\ref{chapt:num}.
