\vssub
\subsection{~Selecting model options} \label{sec:switches}
\vssub

The file {\file switch} in the {\file bin} directory contains a set of strings
identifying model options to be selected. Many options are available. Of
several groups of options it is mandatory to select exactly one. These
mandatory switches are described in \para\ref{sub:man_switch}. Other switches
are optional, and are described in \para\ref{sub:opt_switch}. Default model
setting are identified in \para\ref{sub:opt_default}. The order in which the
switches appear in {\file switch} is arbitrary. How these switches are
included in the source code files is described in \para\ref{sec:w3adc}.

\vsssub
\subsubsection{~Mandatory switches} \label{sub:man_switch}
\vsssub

Of each of the below groups of switches exactly one has to be selected. The
first group of switches controls the selection of machine-dependent code. With
the introduction of \fortran-90 this set of switches should have become
obsolete. Problems with some compilers have prompted the retention of the
second switch.
\begin{slist}
\sit{f90} {\fortran-90 style date and time capturing and program
           abort.}
\sit{dum} {Dummy to be used if \ws\ is to be installed on
           previously untried hardware.}
\end{slist}

\noindent
Hardware model (first group) and message passing protocol (second group). Note
that these two groups share a switch. This implies that the {\sc mpi} switch
can only be used in combination with the {\sc dist} switch.
\begin{slist}
\sit{shrd}{Shared memory model.}
\sit{dist}{Distributed memory model}.
\end{slist}

\begin{slist}
\sit{shrd}{Shared memory model, no message passing.}
\sit{mpi} {Message Passing Interface (MPI).}
\end{slist}

\noindent
Selection of propagation schemes and GSE alleviation method. These represent
two sets of switches with some shared switches between the groups. Note that
the second set of switches is secondary to the selection of program modules
in the first set of switches, and therefore, does not have a user-defined
option.
\begin{slist}
\sit{pr0} {No propagation scheme / GSE alleviation used.}
\sit{pr1} {First order propagation scheme, no GSE alleviation.}
\sit{pr2} {Higher-order schemes with \cite{art:BH87} dispersion correction.}
\sit{pr3} {Higher-order schemes with \cite{tol:OMOD02b} averaging technique.}
%\sit{pr4} {\uq\ propagation scheme with \cite{tol:OMOD02b}
%           divergence technique.}
\sit{prx} {Experimental (user supplied).}
\end{slist}

\begin{slist}
\sit{pr0} {No propagation scheme used.}
\sit{pr1} {First-order propagation scheme.}
\sit{uno} {Second-order (UNO) propagation scheme.}
\sit{uq } {Third-order (UQ) propagation scheme.}
\end{slist}

\noindent
Selection of flux computation:
\begin{slist}
\sit{flx0} {No routine used; flux computation included in source terms,}
\sit{flx1} {Friction velocity according to Eq.~(\ref{eq:Wu}).}
\sit{flx2} {Friction velocity from Tolman and Chalikov input.}
\sit{flx3} {Idem, with cap of Eq.~(\ref{eq:Cd_cap_1}) or (\ref{eq:Cd_cap_2}).}
\sit{flx4} {Friction velocity according to Eq.~(\ref{eq:ST607}).}
\sit{flxx} {Experimental (user supplied).}
\end{slist}

\noindent
Selection of linear input:
\begin{slist}
\sit{ln0} {No linear input.}
\sit{seed}{Spectral seeding of Eq.~(\ref{eq:seed}).}
\sit{ln1} {Cavaleri and Malanotte-Rizzoli with filter.}
\sit{lnx} {Experimental (user supplied).}
\end{slist}

\noindent
Selection of input and dissipation. {\F stab{}\it n} switches are optional and
additional to corresponding {\F st{\it n}} switch:
\begin{slist}
\sit{st0} {No input and dissipation used.}
\sit{st1} {\wam\-3 source term package.}
\sit{st2} {\cite{tol:JPO96} source term package. See also the optional 
          {\F stab2} switch.}
\sit{stab2}{Enable stability correction (\ref{eq:scor}) - (\ref{eq:stab})
            for {\F st2}.}
\sit{st3} {\wam\-4 and variants source term package.}
\sit{stab3}{Enable stability correction from \cite{rep:AB02} for {\F st4}.}
\sit{st4} {\cite{art:Aea10} source term package.}
%\sit{st5} {UNSW source term package.}
\sit{st6} {BYDRZ source term package.}
\sit{stx} {Experimental (user supplied).}
\end{slist}

\noindent
Selection of nonlinear interactions:
\begin{slist}
\sit{nl0} {No nonlinear interactions used.}
\sit{nl1} {Discrete interaction approximation (\dia).}
\sit{nl2} {Exact interaction approximation (\xnl).}
\sit{nl3} {Generalized Multiple \dia\ (\gmd).}
\sit{nl4} {Two-scale approximation (TSA).} 
\sit{nlx} {Experimental (user supplied).}
\end{slist}

\noindent
Selection of bottom friction:
\begin{slist}
\sit{bt0} {No bottom friction used.}
\sit{bt1} {\js\ bottom friction formulation.}
%\sit{bt2} {??? bottom friction formulation.}
%\sit{bt3} {??? bottom friction formulation.}
\sit{bt4} {\showex\ bottom friction formulation.}
\sit{bt8} {Dalrymple and Liu formulation (fluid mud seafloor).}
\sit{bt9} {Ng formulation (fluid mud seafloor).}
\sit{btx} {Experimental (user supplied).}
\end{slist}

\noindent
Selection of term for damping by sea ice:
\begin{slist}
\sit{ic0} {No damping by sea ice.}
\sit{ic1} {Simple formulation.}
\sit{ic2} {Liu et al. formulation.}
\sit{ic3} {Wang and Shen formulation.}
\sit{ic4} {Frequency-dependent damping by sea ice.}
\end{slist}

\noindent
Selection of term for scattering by sea ice:
\begin{slist}
\sit{is0} {No scattering by sea ice.}
\sit{is1} {Diffusive scattering by sea ice (simple).}
\sit{is2} {Floe-size dependent scattering and dissipation.}
\end{slist}

\noindent
Selection of term for reflection:
\begin{slist}
\sit{ref0} {No reflection.}
\sit{ref1} {Enables reflection of shorelines and icebergs}
\end{slist}

\noindent
Selection depth-induced breaking of :
\begin{slist}
\sit{db0} {No depth-induced breaking used.}
\sit{db1} {Battjes-Janssen.}
\sit{dbx} {Experimental (user supplied).}
\end{slist}

\noindent
Selection of triad interactions:
\begin{slist}
\sit{tr0} {No triad interactions used.}
\sit{tr1} {Lumped Triad Interaction (LTA) method.}
\sit{trx} {Experimental (user supplied).}
\end{slist}

\noindent
Selection of bottom scattering:
\begin{slist}
\sit{bs0} {No bottom scattering used.}
\sit{bs1} {Magne and Ardhuin.}
\sit{bsx} {Experimental (user supplied).}
\end{slist}

\noindent
Selection of supplemental source term:
\begin{slist}
\sit{xx0} {No supplemental source term used.}
\sit{xxx} {Experimental (user supplied).}
\end{slist}

\noindent
Selection of method of wind interpolation (time):
\begin{slist}
\sit{wnt0}{No interpolation.}
\sit{wnt1}{Linear interpolation.}
\sit{wnt2}{Approximately quadratic interpolation.}
\end{slist}

\noindent
Selection of method of wind interpolation (space):
\begin{slist}
\sit{wnx0}{Vector interpolation.}
\sit{wnx1}{Approximately linear speed interpolation.}
\sit{wnx2}{Approximately quadratic speed interpolation.}
\end{slist}

\noindent
Selection of method of current interpolation (time):
\begin{slist}
\sit{crt0}{No interpolation.}
\sit{crt1}{Linear interpolation.}
\sit{crt2}{Approximately quadratic interpolation.}
\end{slist}

\noindent
Selection of method of current interpolation (space):
\begin{slist}
\sit{crx0}{Vector interpolation}
\sit{crx1}{Approximate linear speed interpolation.}
\sit{crx2}{Approximate quadratic speed interpolation.}
\end{slist}

\noindent
Switch for user supplied GRIB package.
\begin{slist}
\sit{nogrb}{No package included.}
\sit{ncep1}{\ncep\ GRIB1 package for IBM SP.}
\sit{ncep2}{\ncep\ GRIB2 package for IBM SP.}
\end{slist}

% -------------------------------------------------------------
\vsssub
\subsubsection{~Optional switches} \label{sub:opt_switch}
\vsssub

All switches below activate model behavior if selected, but do not require
particular combinations. The following switches control optional output for
\ws\ programs.

\begin{slist}
\sit{o0}  {Output of namelists in grid preprocessor.}
\sit{o1}  {Output of boundary points in grid preprocessor.}
\sit{o2}  {Output of the grid point status map in grid preprocessor.} 
\sita{o2} {Generation of land-sea mask file {\file mask.ww3} in grid
           preprocessor.}
\sitb{o2} {Output of obstruction map in grid preprocessor.}
\sitc{o2} {Print status map in format as read by {\file ww3\_grid}.}
\sit{o3}  {Additional output in loop over fields in field preprocessor.}
\sit{o4}  {Print plot of normalized one-dimensional energy
           spectrum in initial conditions program.}
\sit{o5}  {Id. two-dimensional energy spectrum.}
\sit{o6}  {Id. spatial distribution of wave heights (not adapted for 
           distributed memory).}
\sit{o7}  {Echo input data for homogeneous fields in generic shell.}
\sita{o7} {Diagnostic output for output points.}
\sitb{o7} {Idem in {\file ww3\_multi}.}
\sit{o8}  {Filter field output for extremely small wave heights
           in wave model (useful for some propagation tests).}
\sit{o9}  {Assign a negative wave height to negative energy in wave model.
           Used in testing phase of new propagation schemes.}
\sit{o10} {Identify main elements of multi-grid model extensions in
           standard output.}
\sit{o11} {Additional log output on management algorithm in {\file log.mww3}.}
\sit{o12} {Identify removed boundary points in overlapping grids (center).}
\sit{o13} {Identify removed boundary points in overlapping grids (edge).}
\sit{o14} {Generate log file with buoy data {\file buoy\_log.ww3} for output
           type {\code ITYPE = 0} in {\file ww3\_outp}.}
\sit{o15} {Generate log file with time stamps of input data file {\file
           times.XXX} in {\file ww3\_prep}.}
\sit{o16} {Generate GrADS output of grid partitioning in {\file ww3\_gspl}.}
\end{slist}

\noindent
The following switches enable parallelization of the model using \omp\
directives, also known as `threading'. Before model version 5.01, threading
and parallelization using the {\sc mpi} switch could no be used
simultaneously. With version 5.01, pure MPI,pure OMP and hybrid MPI-OMP
approaches became available. Switches used in version 5.01 and higher are not
compatible with switches used in previous model versions.
\begin{slist}
\sit{ompg}{General loop parallelization directives used for both exclusive
    OpenMP parallelization and hybrid MPI-OpenMP parallelization.}
\sit{ompx}{Idem, but for directives used only for exclusive OpenMP
    parallelization.}  
\sit{omph}{Idem, but for directives used only for hybrid MPI-OpenMP
    parallelization.}
\end{slist}
Note that these switches can only be used in certain combinations, as enforced
in the model installation scripts (particularly {\file make\_makefile.sh}. A
pure MPI approach requires the {\F dist} and {\F mpi} switches. A pure OpenMP
approach requires the {\F shrd}, {\F ompg} and {\F ompx} switches, and the
hybrid approach requires the {\F dist}, {\F mpi}, {\F ompg}, and {\F omph}
switches.

\vspace{\baselineskip}
\noindent
The following switches are associated with the continuously moving grid
options. The first switch activates the option, the other two are optional
additions.
\begin{slist}
\sit{mgp }{Activate propagation correction in
           Eq.~(\ref{eq:bal_move}).} 
\sit{mgw }{Apply wind correction in moving grid approach.}
\sit{mgg }{Activate GSE alleviation correction in
           Eq.~(\ref{eq:move_GSE_avg2}).}
\end{slist}

\noindent
The following compiler dependent switches are available. They may not have
been maintained for recent compiler versions.
\begin{slist}
\sit{c90} {Compiler directives for Cray C90 (vectorization).}
\sit{nec} {Compiler directives for NEC SX6/SX8 (vectorization).}
\end{slist}

\noindent
Furthermore the following miscellaneous switches are available:
\begin{slist}
\sit{arc} {Arctic grid option for SMC grid\footnote{~Not yet fully tested according to author.}.}
\sit{cou} {Activates the calculation of variables required for coupling}
\sit{dss0}{Switch off frequency dispersion in diffusive
           dispersion correction.}
\sit{fld1}{Sea-state dependent $\tau$ \citet{art:Rei14} (\para\ref{sec:FLD1}).}      
\sit{fld2}{Sea-state dependent $\tau$ \citet{art:Don12} (\para\ref{sec:FLD2}).}
\sit{ig1} {Second-order spectrum and free infragravity waves (\para\ref{sec:IG1}.}        
\sit{mlim}{Use Miche-style shallow water limiter of Eq.~(\ref{eq:MLIM})}.
\sit{mpibdi}{Experimental parallelization of multi-grid model initialization.}
\sit{mpit}{Test output for \mpi\ initializations.}
\sit{mprf}{Profiling of individual models and nesting in {\file ww3\_multi}.}
\sit{nc4} {Activates the NetCDF-4 API in the NetCDF pre- and post-processing
           programs.}
\sit{ncc} {NCEP coupler.}
\sit{nco} {Code modifications for operational implementation at NCO
           (NCEP Central Operations). Mostly changes unit numbers
           and file names. Not recommended for general use.} 
\sit{nls }{Activate nonlinear smoother (\para\ref{sec:NLS}).}
\sit{nnt} {Generate file test\_data\_{\it{nnn}}.ww3 with spectra and 
           nonlinear interactions for training and testing of NNIA}.
\sit{oasis }{Initializes OASIS Coupler (App.~\ref{sec:couplingB}).}   
\sit{oasacm}{OASIS atmospheric model coupling fields(App.~\ref{sec:couplingB}).}
\sit{oasocm}{OASIS oceanic model coupling fields (App.~\ref{sec:couplingB}).}
\sit{refrx} {Enables refraction based on spatial gradients in phase velocity (\para\ref{sec:ICE3})}
\sit{reft}{Test output for shoreline reflection (which is activated with
          {\F ref1}).}
\sit{rtd} {Rotated grid option.}
\sit{rwnd}{Correct wind speed for current velocity.}
\sit{s}   {Enable subroutine tracing in the main \ws\
           subroutines by activating calls to the
           subroutine {\F strace}.}
\sit{scrip   }{Enable SCRIP remapping routines (App. \ref{sec:scripC})}
\sit{scripnc }{Enable storage of remapping weights in NetCDF files
 (App. \ref{sec:scripC})}
\sit{sec1}{Enable the use of global time steps less than 1~s, but does not
           allow output at time steps less than 1~s.}
\sit{smc} {Activate SMC grid.}
\sit{t}   {Enable test output throughout the program(s).}
\sitn{t}  {Id.}
\sit{tdyn}{Dynamic increment of swell age in diffusive
           dispersion correction (test cases only).}
\sit{tide }{Enables tidal analysis: used for pre-processing of input
           files, run-time tidal prediction in ww3\_shel or tidal prediction
           with ww3\_prtide.}
\sit{tidet}{test output for tidal analysis.}
\sit{trknc} {Activates the NetCDF API in the wave system tracking
           post-processing program. Selecting TRKNC alone will generate 
           NetCDF-3 files. Selecting both TRKNC and NC4 will generate 
           NetCDF-4 files.}
\sit{uost}{Enable the unresolved obstacles source term.}
\sit{xw0 }{Swell diffusion only in \uq\ scheme.}
\sit{xw1 }{Id. wave growth diffusion only.}
\end{slist}

\vsssub
\subsubsection{~Default model settings} \label{sub:opt_default}
\vsssub

Up to model version 3.14, the NCEP operational model setup was considered as
the default model setup. However, with subsequent versions of \ws, the model
has evolved into a modeling framework rather than a single model. With this,
\ws\ is run differently at various centers, and a clear ``default'' model
version can no longer be identified.  Nevertheless, in order to be able to
concisely identify in publications exactly which model setup is used,
``default'' configurations of various centers are now provided in the {\file
bin} directory. These configurations are provided in example switch files and
README files, such as {\file switch\_NCEP\_st2} and {\file README.NCEP}. Note
that these files are provided to simplify referring to model version, but do
not imply an endorsement of the specific model configuration.; in this
context, it should be noted that by nature, model versions at operational
centers are in a continuous state of development.



% With the plethora of numerical and physical options in \ws, and with the different and ever evolving way \ws\ is implemented at various organizations, we no longer provide a `default' model set up for the model with respect to most numerical and physical approaches. The only setup options that are still considered default is the treatment of input fields.

% \begin{slist}
% \sit{wnt1} {Wind interpolation (linear).}
% \sit{wnx1} {Wind interpolation (linear).}
% \sit{rwnd} {Define wind as relative to current.}
% \sit{crt1} {Current interpolation (linear).}
% \sit{crx1} {Current interpolation (linear).}
% \end{slist}
