\section{~Introduction}
\vssub
\subsection{~About this manual}
\vssub

This is the user manual and system documentation of version \WWver\ of the
full-spectral third-generation wind-wave model \ww. This model has been
developed at the Marine Modeling and Analysis Branch (MMAB) of the
Environmental Modeling Center (EMC) of the National Centers for Environmental
Prediction (NCEP). It is based on WAVEWATCH~I and WAVEWATCH~II as developed at
Delft University of Technology, and NASA Goddard Space Flight Center,
respectively. \ws\ differs from its predecessors in all major aspects; i.e.,
governing equations, program structure, numerical and physical
approaches. %\wt\ is a trademark of the United States government.

This manual describes the governing equations, numerical approaches,
compilation, and running of \ws. The format of a combined user manual and
system documentation has been chosen, to give users the necessary background
to upgrade the model according to their own specifications.  This approach is
becoming more important, as \ws\ is developing into a wave modeling framework
rather than a wave model. By design, a user can apply his or her numerical or
physical approaches, and thus develop a new wave model based on the \ws\
framework. In such an approach, optimization, parallelization, nesting, input
and output service programs from the framework can be easily shared between
actual models.  Whereas this document is intended to be complete and
self-contained, this is not the case for all elements in the system
documentation. For additional system details, reference is made to the source
code, which is fully documented.

The governing equations and numerical approaches used in this model are
described in chapters~\ref{chapt:eq} and \ref{chapt:num}. Running the model is
described in chapter~\ref{chapt:run}. Installing \ws\ is described in
chapter~\ref{chapt:impl}. Finally, a short system documentation is given in
chapter~\ref{chapt:sys}. A thorough knowledge of \ws\ can be obtained by
following chapters~\ref{chapt:eq} through \ref{chapt:impl}. A shortcut is to
first install the model (chapter~\ref{chapt:impl}), and then successively
modify input files in example runs (chapter~\ref{chapt:run}).

\vspace{\baselineskip} \pb \noindent The present model version (\WWver) is the
first official release since model version 2.22. Since the latter release
massive modifications have been made to the code.  Some of the major
modifications are:

\begin{list}{$\bullet$}{\rightmargin 5mm \parsep 0mm \itemsep 0mm}

\item
Development of the multi-grid or wave model driver {\file ww3\_multi}. In this
driver, the full two-way interaction between all grids is considered at the
time step level. The development of this approach is described in
\cite{tol:MMAB07b, tol:OMOD08b}. Idealized test cases {\file mww3\_test\_01}
through {\file 05}, and real life test cases {\file mww3\_case\_01} through
{\file 03} have been added to the standard test cases to demonstrate the
capabilities of this approach.

\item
Source term slots for linear wind input and surf-zone physics have been added.
New source term packages have been included such as linear input, surf
breaking, \wam\-4 input and dissipation and bottom scattering. The exact
nonlinear interactions routines (WRT) have been upgraded to version 5 of
this package.

\item
To make the development of a multi-grid model possible, a new dynamic data
structure was introduced to the model. This has little or no impact on the
model user, but it has major consequences for model developers.

\item
A model version with a continuously moving grid has been developed
\citep[model version 3.02,][]{tol:OMOD05b}. This model version is intended
only for deep water without land masses or currents. As part of this
implementation, the calculation of advection time steps in Eq.~(\ref{eq:dtpl})
has been revised. To test this new option, a hurricane test {\file ww3\_ts3}
has been added to the standard test cases.

\item
In model version 3.04 the track output file format is converted to a
sequential file, using a single file server process, and file server options
are introduced for the restart files \citep{tol:MMAB03b}. In later model
version, the option is added to designate some processes to perform output
only.

\item
Model version 3.05 includes expanded GrADS scripts to plot spectra and source
terms, adding black and white options, and an alternative Cartesian display
format.

\item
Wind stresses in the model are now described as vectors, paving the way for
flux parameterizations that allow for stresses that do not line up with the
wind direction. Additional output fields are added. Partitioning of spectral
data can now be done for the entire wave model grid (new output type), and
provides partitioned wind sea and swell output fields (publication in
preparation).

\item
With the introduction of the multi-grid model driver, active boundary points
can now be placed at the outer points of the grid. Also, grid points can now
be defined as inactive rather than land, and an expanded definition of the
grid status map allows for drying out and wetting of model grid points.

\item
In the multi-grid model driver, changes of spectral resolution between grids
are now allowed.

\item
Time steps can now be fractions of seconds, although general time management
still propagates the solution by integer seconds.

\end{list}

\vspace{\baselineskip} \noindent 
A note of caution. \ww\ includes wave-current interactions. The implementation
of these interactions has only been tested in idealized test cases. Only
limited tests in realistic conditions have been performed.

\vspace{\baselineskip} \noindent 
Up to date information on this model can be found (including bugs and bug
fixes) on the \ww\ web page, and comments, questions and suggestions should be
directed to the corresponding E-mail address

\begin{center}
http://polar.ncep.noaa.gov/waves/wavewatch \\
NCEP.EMC.wavewatch@NOAA.gov
\end{center}

\noindent
We will redirect questions regarding contributions from outside NCEP to the
respective authors of the codes.


\vssub
\subsection{~Licensing terms}
\vssub

Starting with model version 3.14, \ww\ is distributed under the following
licensing terms:

\vspace \baselineskip \noindent
\centerline{
\rule[1mm]{47mm}{.5mm} {\rm start of licensing terms}
\rule[1mm]{47mm}{.5mm}}

\vspace \baselineskip \noindent 
Software, as understood herein, shall be broadly interpreted as being
inclusive of algorithms, source code, object code, data bases and related
documentation, all of which shall be furnished free of charge to the Licensee.

Corrections, upgrades or enhancements may be furnished and, if furnished,
shall also be furnished to the Licensee without charge. NOAA, however, is not
required to develop or furnish such corrections, upgrades or enhancements.

NOAA's software, whether that initially furnished or corrections or upgrades,
are furnished “as is.” NOAA furnishes its software without any warranty
whatsoever and is not responsible for any direct, indirect or consequential
damages that may be incurred by the Licensee.  Warranties of merchantability,
fitness for any particular purpose, title, and non-infringement, are
specifically negated.

The Licensee is not required to develop any software related to the licensed
software.  However, in the event that the Licensee does so, the Licensee is
required to offer same to NOAA for inclusion under the instant licensing terms
with NOAA's licensed software along with documentation regarding its
principles, use and its advantages.  This includes changes to the wave model
proper including numerical and physical approaches to wave modeling, and
boundary layer parameterizations embedded in the wave model The Licensee is
encouraged but not obligated to provide pre-and post processing tools for
model input and output.  The software required to be offered shall not include
additional models to which the wave model may be coupled, such as oceanic or
atmospheric circulation models. The software provided by the Licensee shall be
consistent with the latest model version available to the Licensee, and
interface routines to the software provided shall conform to programming
standards as outlined in the model documentation. The software offered to NOAA
shall be offered as is, without any warranties whatsoever and without any
liability for damages whatsoever.  NOAA shall not be required to include a
Licensee's software as part of its software.  Licensee's offered software
shall not include software developed by others.

A Licensee may reproduce sufficient software to satisfy its needs.  All copies
shall bear the name of the software with any version number as well as
replicas of any applied copyright notice, trademark notice, other notices and
credit lines.  Additionally, if the copies have been modified, e.g. with
deletions or additions, this shall be so stated and identified.

All of Licensee's employees who have a need to use the software may have
access to the software but only after reading the instant license and stating,
in writing, that they have read and understood the license and have agreed to
its terms.  Licensee is responsible for employing reasonable efforts to assure
that only those of its employees that should have access to the software, in
fact, have access.

The Licensee may use the software for any purpose relating to sea state
prediction.

No disclosure of any portion of the software, whether by means of a media or
verbally, may be made to any third party by the Licensee or the Licensee's
employees

The Licensee is responsible for compliance with any applicable export or
import control laws of the United States.

\vspace \baselineskip \noindent
\centerline{
\rule[1mm]{48mm}{.5mm} {\rm end of licensing terms}
\rule[1mm]{48mm}{.5mm}}

\vspace \baselineskip \noindent
The software will be distributed through our web site after the Licensee has
agreed to the license terms.


\vssub
\subsection{~Copyrights and trademarks}
\vssub

\ww\ \copyright\ 2009 National Weather Service, National Oceanic and
Atmospheric Administration.  All rights reserved. \ww\ is a trademark of the
National Weather Service. No unauthorized use without permission.

\vssub
\subsection{~Contributors and acknowledgments}
\vssub

Even in its original development, when I was working on the code mostly on my
own on the development of \ww, many have contributed to the success of the
model. With the expansion of physical and numerical parameterizations
available, the list of contributors to this model is growing. I would like to
recognize the following contributors (in alphabetic order) :

\begin{list}{\ldots}{ }
\item [Henrique Alves] (Metocean Engineers, Australia) \\
Support of code development while at NCEP, shallow water physics packages.
\item [Fabrice Ardhuin] (Service Hydrographique et Oc\'{e}anographique de la
Marine, France) \\
Various  physics packages.
\item [Nico Booij] (Delft University of Technology, The Netherlands) \\
Original design of source code pre-processor ({\code w3adc}), basic method of
documentation and other programming habits. Spatially varying wavenumber grid.
\item [Dmitry V. Chalikov] (ESSIC, Univ. of Maryland, USA) \\ Co-author of the
\cite{tol:JPO96} input and dissipation parameterizations and source code.
\item [Arun Chawla](Science Applications International Corporation, USA) \\
Support of code development at NCEP, GRIB packing, automated grid generation
software \citep{tol:MMAB07a, tol:OMOD08a}.
%\item [Vladimir Krasnopolsky] (Science Applications International Corporation,
%  USA) \\
%Neural Network nonlinear interaction approaches.
\item [Barbara Tracy] (US Army Corps of Engineers, ERDC-CHL, USA) \\
Spectral partitioning
\item [Gerbrant Ph. van Vledder] (Alkyon Hydraulic Consultancy \& Research,
NL) \\ 
Webb-Resio-Tracy exact nonlinear interaction routines, as well as some of the
original service routines.
\end{list}

\noindent
The development of \ww\ has been an ongoing process for well over a
decade. The development of WAVEWATCH I was entirely funded through my
Ph.D. work at Delft University. The development of WAVEWATCH II has been
funded entirely through my position as a National Research Council Resident
Research Associate at NASA, Goddard Space Flight Center. The initial
development of WAVEWATCH III version 1.18 was entirely funded by NOAA/NCEP,
with most funding provided by the NOAA High Performance Computing and
Communication (HPCC) office. Developments of version 2.22 have been funded
similarly through NOAA/NCEP. Since then, funding is provided by NCEP, but also
by many partners outside NCEP. Much of the work at NCEP has been performed
under contract by SAIC.

I would finally like to thank all users who have reported errors and glitches,
or have made suggestions for improvements, particularly those who have
beta-tested this and previous model release.

%\bpage
