\vsssub
\subsubsection{Gridded GRIB output post-processor} \label{sec:ww3grib}
\vsssub

\proddefH{ww3\_grib}{w3grib}{ww3\_grib.ftn}
\proddeff{Input}{ww3\_grib.inp}{Input file for gridded output
post-processor.}{10}
\proddefa{mod\_def.ww3}{Model definition file.}{20}
\proddefa{out\_grd.ww3}{Raw gridded output data.}{20}
\proddeff{Output}{standard out}{Formatted output of program.}{6}
\proddefa{gribfile}{GRIB file.}{50}

\inpfile{ww3_grib.tex}

\vspace{\baselineskip} 
\noindent
This post-processor packs fields of mean wave parameters in GRIB format, using
GRIB version II and \ncep's w3 and bacio library routines, or in GRIB2, using
NCEPS's operational package. Additional packing data can be found in
Table~\ref{tab:fields} on page \pageref{tab:fields}.

The GRIB packing is performed using the \ncep's GRIB tables as described in
\cite{rep:GRIB1}. Because the w3 and bacio routine are not fully portable, they
are not supplied with the code. The user will have to provide corresponding
routines. It is suggested that such routines are activated with additional
\ws\ switches in the mandatory switch group containing the `{\F nogrb}'
switch, as if presently the case with the \ncep\ routines.  The GRIB2 packing
is performed according to \cite{rep:GRIB2}, and is performed with NCEP's
standard operational packages.

Table~\ref{tab:fields} shows the {\F kpds(5)} data values for GRIB
packing. For the partitioned data, the first number identifies the wind sea,
the second number identifies swell. Most data are packed as surface data ({\F
kpds(6) = 0}). For the partitioned swell fields, however, consecutive fields
are packed at consecutive levels, with the level type indicator set to ({\F
kpds(6) = 241}). {\F kpds(7)} identifies the actual level or swell field
number.

Table~\ref{tab:fields} shows several {\F kpds} data values for GRIB2
packing. The first number in the table represents {\F listsec0(2)}, which
identifies the discipline type (e.g., oceanography, meteorology, etc.)  The
second number represents {\F kpds(1)}, which identifies the parameter category
(e.g., waves, circulation, ice, etc.) within the discipline type.  The third
number represents {\F kpds(2)}, which identifies the actual parameter.  For
the partitioned data, A/B means A for wind sea and B for swell.  Additionally
{\F kpds(10) = 0} for surface data, and {\F kpds(10) = 241 } to pack
consecutive swell fields at consecutive levels. {\F kpds(12)} identifies the
actual level or swell field number.

Although the above input file contains flags for all 31 output fields of \ws,
not all fields can be packed in GRIB. If a parameter is chosen for which GRIB
packing is not available, a message will be printed to standard
output. Table~\ref{tab:fields} shows which parameter can be packed in GRIB.
Note that at \ncep\ the conversions from GRIB to GRIB2 coincided with the
introduction of partitioned wave model output. This required some duplicate
definitions in GRIB and some apparent inconsistencies between GRIB and GRIB2
packing.

\pb